---
title: "Spatial Analytics Analysis"
author: "Shiraz Ben Shoshan"
date: "2025-05-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = 'set working directory here')
```

## this analysis starts by creating a shiny app for paris landmarks and hotels and then creates the same for copenhagen landmarks and hotels

```{r}
# # installing packages - make sure to install these if needed
# install.packages("sf")
# install.packages("osmdata")
# install.packages("leaflet")
# install.packages("units")
# install.packages("tidyverse")
# install.packages("mapboxapi")
#install.packages("geosphere")
#install.packages("rnaturalearth")
```

```{r}
# loading the packages
library(sf)
library(osmdata)
library(tidyverse)
library(leaflet)
library(units)
library(dplyr)
library(mapboxapi)
library(shiny)
library(readr)
library(geosphere)
library(ggplot2)
library(scales)
library(rnaturalearth)
```

# getting the paris landmarks ready
```{r}
# loading in pre-made geojson file with paris landmarks
paris_landmarks <- st_read("paris-landmarks.geojson")

# transforming the data to common CRS
paris_landmarks <- st_transform(paris_landmarks, 3857)
```

```{r}
# cleaning up paris landmarks df for later use 

# transformation to CRS needed for mapboxapi
paris_landmarks <- st_transform(paris_landmarks, 4326)

# overwriting the df to include only the geometry
paris_landmarks <- paris_landmarks %>% 
  select(geometry)

# finding central point for each landmark for analysis
paris_landmark_points <- st_centroid(paris_landmarks)

# printing to examine
paris_landmark_points
```

```{r}
# creating a list with the landmark names for later analysis
paris_landmark_names <- c("Notre Dame Cathedral", "Jardin du Luxembourg", "Sacre-Coeur", "Jardin des Tuileries", "Louvre Museum",
                    "Eiffel Tower", "Arc de Triumphe", "Palais Royal", "Pantheon", "Orsay Museum", "Palais Garnier",
                    "Pont Alexandre III", "Les Invalids", "Champs-Elysees", "Place de la Concorde", "Sainte-Chapelle",
                    "Moulin Rouge", "Pompidou Center", "Rodin Museum", "Pont Neuf")

# overwriting the df to include landmark names
paris_landmark_points <- paris_landmark_points %>% 
  mutate(landmark_name = paris_landmark_names)
```

# getting the paris hotels ready
```{r}
# using OpenStreetMap for collecting hotels

# define bounding box for Paris 
bbox <- c(
  min = 2.252,  # Western Paris
  minlat = 48.816,
  max = 2.422,  # Eastern Paris
  maxlat = 48.902
)

# code to bypass internet issues
assign("has_internet_via_proxy", TRUE, environment(curl::has_internet))

# query hotels using Overpass API
hotels_query <- opq(bbox = bbox) %>%
  add_osm_feature(key = "tourism", value = "hotel") %>%
  osmdata_sf()

# extract hotel points
paris_hotels <- hotels_query$osm_points %>%
  st_transform(3857)  

# clean missing geometry
paris_hotels <- paris_hotels[!st_is_empty(paris_hotels), ]

# cleaning out df to disregard NA values and ensure name of hotel is seen as character
paris_hotels <- paris_hotels %>%
  filter(!is.na(name)) %>%
  mutate(label_name = as.character(name))
```

```{r}
# cleaning the df to only contain necessary info and ensure long and lat coordinates are set up correctly
paris_hotels_coords <- paris_hotels %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) %>%
  dplyr::select(name, osm_id, lon, lat)

paris_hotels_coords
```

```{r}
# setting seed to ensure replicability
set.seed(123)  

# sample 100 random rows from paris_hotels_coords since the df is too big (includes over 1,200 hotels)
hotels_sample_100 <- paris_hotels_coords %>%
  slice_sample(n = 100)
```

# calculating distance and walking time to between every hotel in the sample to every one of the defined 20 landmarks using mapboxapi

```{r}
# set token
Sys.setenv(MAPBOX_ACCESS_TOKEN = "your access token goes here")

# transform to WGS84
paris_hotels_wgs <- st_transform(hotels_sample_100, 4326)
paris_landmark_points_wgs <- st_transform(paris_landmark_points, 4326)

# extract coords + names from transformed data
final_paris_hotels <- paris_hotels_wgs %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) %>%
  st_drop_geometry() %>%
  select(name, lon, lat)

# prepare coordinate matrices
origins <- as.matrix(final_paris_hotels[, c("lon", "lat")])
destinations <- as.matrix(st_coordinates(paris_landmark_points_wgs))

# check that matrices are set up correctly
stopifnot(ncol(origins) == 2, ncol(destinations) == 2)

# convert to list of [lon, lat] pairs
origins_list <- split(origins, seq(nrow(origins)))
origins_list <- lapply(origins_list, as.numeric)

destinations_list <- split(destinations, seq(nrow(destinations)))
destinations_list <- lapply(destinations_list, as.numeric)

# chunk size (5 origins per chunk to stay under 25 coordinate limit)
chunk_size <- 5
chunks <- split(origins_list, ceiling(seq_along(origins_list) / chunk_size))

# initialize lists to collect results
distances_list <- list()
durations_list <- list()

# loop through chunks to call Mapbox API safely
for (i in seq_along(chunks)) {
  cat("Processing chunk", i, "of", length(chunks), "\n")
  
  dist_chunk <- mb_matrix(
    origins = chunks[[i]],
    destinations = destinations_list,
    profile = "walking",
    access_token = Sys.getenv("MAPBOX_ACCESS_TOKEN"),
    output = "distance"
  )
  
  dur_chunk <- mb_matrix(
    origins = chunks[[i]],
    destinations = destinations_list,
    profile = "walking",
    access_token = Sys.getenv("MAPBOX_ACCESS_TOKEN"),
    output = "duration"
  )
  
  distances_list[[i]] <- dist_chunk
  durations_list[[i]] <- dur_chunk
}

# combine all chunks into full matrices
distances <- do.call(rbind, distances_list)
durations <- do.call(rbind, durations_list)

# build result table
paris_hotel_landmark <- expand.grid(
  hotel = final_paris_hotels$name,
  landmark = paris_landmark_points$landmark_name
)

paris_hotel_landmark$distance_m <- as.vector(distances)
paris_hotel_landmark$duration_m <- as.vector(durations)

# view sample of result
head(paris_hotel_landmark)
```

# saving data needed for shiny app

```{r}
write.csv(paris_hotel_landmark, "paris_hotel_landmark.csv", row.names = FALSE)
saveRDS(paris_hotel_landmark, "paris_hotel_landmark.rds")
write.csv(final_paris_hotels, "final_paris_hotels.csv", row.names = FALSE)
saveRDS(paris_landmark_points, "paris_landmark_points.rds")
```

# creating shiny app

```{r}
# load data
hotel_distances <- read_csv("paris_hotel_landmark.csv")
hotel_coords <- read_csv("final_paris_hotels.csv")
landmarks <- readRDS("paris_landmark_points.rds")  
landmarks <- st_as_sf(landmarks)

ui <- fluidPage(
  titlePanel("Paris Hotels Near Landmarks"),
  sidebarLayout(
    sidebarPanel(
      selectizeInput("landmarks", "Select Landmarks:", choices = NULL, multiple = TRUE),
      sliderInput("max_distance", "Max Walking Distance (meters):", min = 0, max = 7000, value = 1000, step = 100)
    ),
    mainPanel(leafletOutput("map", height = 650))
  )
)

server <- function(input, output, session) {
  
  # populate landmark choices
  observe({
    updateSelectizeInput(session, "landmarks",
                         choices = sort(unique(landmarks$landmark_name)),
                         selected = unique(landmarks$landmark_name)[1])
  })
  
  # hotels that are within distance for all selected landmarks
  selected_hotels <- reactive({
    req(input$landmarks)
    
    filtered <- hotel_distances %>%
      filter(landmark %in% input$landmarks, distance_m <= input$max_distance)
    
    hotel_counts <- filtered %>%
      count(hotel) %>%
      filter(n == length(input$landmarks))
    
    hotel_coords %>%
      filter(name %in% hotel_counts$hotel) %>%
      rename(hotel = name)
  })
  
  # tooltip data: Distance and time for each landmark
  hotel_tooltips <- reactive({
    req(input$landmarks)
    
    hotel_distances %>%
      filter(landmark %in% input$landmarks, distance_m <= input$max_distance) %>%
      mutate(time_min = round(duration_m, 1))
  })
  
  # landmark points to show on the map
  selected_landmark_points <- reactive({
    req(input$landmarks)
    landmarks %>% filter(landmark_name %in% input$landmarks)
  })
  
  # initial map
  output$map <- renderLeaflet({
    leaflet() %>%
      addProviderTiles("CartoDB.Positron") %>%
      setView(lng = 2.3522, lat = 48.8566, zoom = 12) %>%
      addLayersControl(overlayGroups = c("Landmark Points", "Hotels"),
                       options = layersControlOptions(collapsed = FALSE))
  })
  
  # add landmark markers
  observe({
    leafletProxy("map") %>%
      clearGroup("Landmark Points") %>%
      addCircleMarkers(
        data = selected_landmark_points(),
        lng = ~st_coordinates(geometry)[,1],
        lat = ~st_coordinates(geometry)[,2],
        radius = 8,
        color = "blue", fillColor = "lightblue", fillOpacity = 0.8,
        group = "Landmark Points",
        label = ~landmark_name
      )
  })
  
  # add hotel markers with distance and time info
  observe({
    leafletProxy("map") %>% clearGroup("Hotels")
    
    hotels <- selected_hotels()
    details <- hotel_tooltips()
    if (is.null(hotels) || nrow(hotels) == 0) return()
    
    hotel_tooltip_data <- hotels %>%
      left_join(details, by = "hotel") %>%
      group_by(hotel, lat, lon) %>%
      summarise(
        popup = paste0(
          "<b>", hotel, "</b><br/>",
          paste0(
            landmark, ": distance ", round(distance_m), " m, time: ", round(duration_m, 1), " min",
            collapse = "<br/>"
          )
        ),
        .groups = "drop"
      )
    
    leafletProxy("map") %>%
      addCircleMarkers(
        data = hotel_tooltip_data,
        lng = ~lon,
        lat = ~lat,
        radius = 5,
        color = "red", fillOpacity = 0.7,
        group = "Hotels",
        label = lapply(hotel_tooltip_data$popup, HTML),
        popup = ~popup
      )
  })
}

shinyApp(ui, server)
```

# repeating workflow for copenhagen landmarks and hotels

```{r}
# loading in copenhagen landmarks 
copenhagen_landmarks <- st_read("copenhagen-landmarks.geojson")  

# transforming the data to common CRS 
copenhagen_landmarks <- st_transform(copenhagen_landmarks, 3857)
```

```{r}
# transformation to CRS needed for mapboxapi
copenhagen_landmarks <- st_transform(copenhagen_landmarks, 4326)

# overwriting the df to include only geometry information
copenhagen_landmarks <- copenhagen_landmarks %>% 
  select(geometry)

# finding central point for each landmark for analysis
cph_landmark_points <- st_centroid(copenhagen_landmarks)

print(cph_landmark_points)
```

# adding landmark names for analysis purposes

```{r}
# adding columns with landmark names for cph 
cph_landmark_names <- c("Tivoli Gardens", "Nyhavn", "Amalienborg", "Strøget", "Rundetaarn", "Rosenborg Slot", "Den Lille Havfrue", "Christiansborg Slot", "Frederiksberg Slot", "Gråbrødretorv", "Kongens Nytorv", "Frederiks Kirke", "Vor Frelsers Kirke", "Ny Carlsberg Glyptotek", "Peblinge Sø", "Kastellet", "Christiania", "Nordhavn", "The Royal Library", "Nationalmuseet")

# overwriting df with the landmark names
cph_landmark_points <- cph_landmark_points %>% 
  mutate(cph_landmark_name = cph_landmark_names)
```

# getting the copenhagen hotels ready

```{r}
# getting cph hotels from open street map

# define bounding box for Copenhagen
bbox <- c(
  min = 12.55,    # Western Copenhagen
  minlat = 55.60, # Southern Copenhagen
  max = 12.65,    # Eastern Copenhagen
  maxlat = 55.70  # Northern Copenhagen
)

assign("has_internet_via_proxy", TRUE, environment(curl::has_internet))

# query hotels using Overpass API
hotels_query <- opq(bbox = bbox) %>%
  add_osm_feature(key = "tourism", value = "hotel") %>%
  osmdata_sf()

# extract hotel points
copenhagen_hotels <- hotels_query$osm_points %>%
  st_transform(3857)  

# clean missing geometry
copenhagen_hotels <- copenhagen_hotels[!st_is_empty(copenhagen_hotels), ]

copenhagen_hotels <- copenhagen_hotels %>%
  filter(!is.na(name)) %>%
  mutate(label_name = as.character(name))
```

```{r}
# cleaning the df and preparing the long lat coordinates accordingly
cph_hotels_coords <- copenhagen_hotels %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) %>%
  dplyr::select(name, osm_id, lon, lat)

cph_hotels_coords
```

# calculating the distance and walking time between each hotel in list and every landmark (20 in total) using mapboxapi

```{r}
# set Mapbox token
Sys.setenv(MAPBOX_ACCESS_TOKEN = "your access token goes here")

# transform to WGS84 (lat/lon for Mapbox)
cph_hotels_wgs <- st_transform(cph_hotels_coords, 4326)
cph_landmarks_wgs <- st_transform(cph_landmark_points, 4326)

# extract coordinates + names (drop geometry to avoid issues)
final_cph_hotels <- cph_hotels_wgs %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) %>%
  st_drop_geometry() %>%
  select(name, lon, lat)

# prepare coordinate matrices
origins <- as.matrix(final_cph_hotels[, c("lon", "lat")])
destinations <- as.matrix(st_coordinates(cph_landmarks_wgs))

# confirm dimensions
stopifnot(ncol(origins) == 2, ncol(destinations) == 2)

# convert to list of [lon, lat] pairs
origins_list <- split(origins, seq(nrow(origins)))
origins_list <- lapply(origins_list, as.numeric)

destinations_list <- split(destinations, seq(nrow(destinations)))
destinations_list <- lapply(destinations_list, as.numeric)

# set chunk size (max 25 coords per Mapbox Matrix request)
chunk_size <- 5
chunks <- split(origins_list, ceiling(seq_along(origins_list) / chunk_size))

# initialize lists to store results
distances_list <- list()
durations_list <- list()

# loop through chunks to get distances and durations
for (i in seq_along(chunks)) {
  cat("Processing chunk", i, "of", length(chunks), "\n")

  dist_chunk <- mb_matrix(
    origins = chunks[[i]],
    destinations = destinations_list,
    profile = "walking",
    access_token = Sys.getenv("MAPBOX_ACCESS_TOKEN"),
    output = "distance"
  )

  dur_chunk <- mb_matrix(
    origins = chunks[[i]],
    destinations = destinations_list,
    profile = "walking",
    access_token = Sys.getenv("MAPBOX_ACCESS_TOKEN"),
    output = "duration"
  )

  distances_list[[i]] <- dist_chunk
  durations_list[[i]] <- dur_chunk
}

# combine chunks into full matrices
distances <- do.call(rbind, distances_list)
durations <- do.call(rbind, durations_list)

# build result table for Copenhagen
cph_hotel_landmark <- expand.grid(
  hotel = final_cph_hotels$name,
  landmark = cph_landmark_points$cph_landmark_name
)

cph_hotel_landmark$distance_m <- as.vector(distances)
cph_hotel_landmark$duration_m <- as.vector(durations)

# view result
head(cph_hotel_landmark)
```

# saving data for shiny app

```{r}
write.csv(cph_hotel_landmark, "cph_hotel_landmark.csv", row.names = FALSE)
saveRDS(cph_hotel_landmark, "cph_hotel_landmark.rds")
saveRDS(cph_landmark_points, "cph_landmark_points.rds")

# extra measures for copenhagen hotel df 

# making sure they are in WGS84
cph_hotels_sf <- st_transform(cph_hotels_coords, 4326)  

# get lat/lon in plain data frame
cph_hotels_coords <- cph_hotels_sf %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) %>%
  st_drop_geometry()

# save df
write.csv(cph_hotels_coords, "cph_hotels_coords.csv", row.names = FALSE)
```

# creating shiny app

```{r}
# load copenhagen data
hotel_distances <- read_csv("cph_hotel_landmark.csv")
hotel_coords <- read_csv("cph_hotels_coords.csv")
landmarks <- readRDS("cph_landmark_points.rds")
landmarks <- st_as_sf(landmarks)

ui <- fluidPage(
  titlePanel("Copenhagen Hotels Near Landmarks"),
  sidebarLayout(
    sidebarPanel(
      selectizeInput("landmarks", "Select Landmarks:", choices = NULL, multiple = TRUE),
      sliderInput("max_distance", "Max Walking Distance (meters):", min = 0, max = 7000, value = 1000, step = 100)
    ),
    mainPanel(leafletOutput("map", height = 650))
  )
)

server <- function(input, output, session) {
  
  # populate landmark choices
  observe({
    updateSelectizeInput(session, "landmarks",
                         choices = sort(unique(landmarks$cph_landmark_name)),
                         selected = unique(landmarks$cph_landmark_name)[1])
  })
  
  # hotels that are within distance for all selected landmarks
  selected_hotels <- reactive({
    req(input$landmarks)
    
    filtered <- hotel_distances %>%
      filter(landmark %in% input$landmarks, distance_m <= input$max_distance)
    
    hotel_counts <- filtered %>%
      count(hotel) %>%
      filter(n == length(input$landmarks))
    
    hotel_coords %>%
      filter(name %in% hotel_counts$hotel) %>%
      rename(hotel = name)
  })
  
  # tooltip data: Distance and time for each landmark
  hotel_tooltips <- reactive({
    req(input$landmarks)
    
    hotel_distances %>%
      filter(landmark %in% input$landmarks, distance_m <= input$max_distance) %>%
      mutate(time_min = round(duration_m, 1))
  })
  
  # landmark points to show on the map
  selected_landmark_points <- reactive({
    req(input$landmarks)
    landmarks %>% filter(cph_landmark_name %in% input$landmarks)
  })
  
  # initial map (centered on Copenhagen)
  output$map <- renderLeaflet({
    leaflet() %>%
      addProviderTiles("CartoDB.Positron") %>%
      setView(lng = 12.5683, lat = 55.6761, zoom = 13) %>%
      addLayersControl(overlayGroups = c("Landmark Points", "Hotels"),
                       options = layersControlOptions(collapsed = FALSE))
  })
  
  # add landmark markers
  observe({
    leafletProxy("map") %>%
      clearGroup("Landmark Points") %>%
      addCircleMarkers(
        data = selected_landmark_points(),
        lng = ~st_coordinates(geometry)[,1],
        lat = ~st_coordinates(geometry)[,2],
        radius = 8,
        color = "blue", fillColor = "lightblue", fillOpacity = 0.8,
        group = "Landmark Points",
        label = ~cph_landmark_name
      )
  })
  
  # add hotel markers with distance and time info
  observe({
    leafletProxy("map") %>% clearGroup("Hotels")
    
    hotels <- selected_hotels()
    details <- hotel_tooltips()
    if (is.null(hotels) || nrow(hotels) == 0) return()
    
    hotel_tooltip_data <- hotels %>%
      left_join(details, by = "hotel") %>%
      group_by(hotel, lat, lon) %>%
      summarise(
        popup = paste0(
          "<b>", hotel, "</b><br/>",
          paste0(
            landmark, ": distance ", round(distance_m), " m, time: ", round(duration_m, 1), " min",
            collapse = "<br/>"
          )
        ),
        .groups = "drop"
      )
    
    leafletProxy("map") %>%
      addCircleMarkers(
        data = hotel_tooltip_data,
        lng = ~lon,
        lat = ~lat,
        radius = 5,
        color = "red", fillOpacity = 0.7,
        group = "Hotels",
        label = lapply(hotel_tooltip_data$popup, HTML),
        popup = ~popup
      )
  })
}

shinyApp(ui, server)
```


# analyses to compare and contrast

# comparing average walking distances to landmarks
```{r}
# copenhagen stats
cph_distance_stats <- cph_hotel_landmark %>%
  summarise(
    mean_distance = mean(distance_m, na.rm = TRUE),
    median_distance = median(distance_m, na.rm = TRUE)
  )

# paris stats
paris_distance_stats <- paris_hotel_landmark %>%
  summarise(
    mean_distance = mean(distance_m, na.rm = TRUE),
    median_distance = median(distance_m, na.rm = TRUE)
  )

# combine for comparison
distance_comparison <- bind_rows(
  cph_distance_stats %>% mutate(city = "Copenhagen"),
  paris_distance_stats %>% mutate(city = "Paris")
)

print(distance_comparison)
```

The analysis reveals that Copenhagen hotels are generally closer to their landmarks compared to Paris. The mean walking distance from hotels to landmarks in Copenhagen is approximately 2,504 meters, while in Paris it is 3,717 meters. Similarly, the median distance is 2,046 meters for Copenhagen and 3,513 meters for Paris. This suggests that Copenhagen offers more centralized access to landmarks, potentially making it easier for tourists to explore the city on foot. In contrast, Paris landmarks appear to be more dispersed or hotels may be more spread out, resulting in longer walking distances overall.

# analyzing landmark spatial spread

```{r}
# extract coordinates in lon-lat for each city
cph_coords <- st_coordinates(st_transform(cph_landmark_points, 4326))
paris_coords <- st_coordinates(st_transform(paris_landmark_points, 4326))

# function to calculate average pairwise distance (meters)
avg_pairwise_dist <- function(coords_matrix) {
  # Calculate all pairwise distances
  dist_matrix <- distm(coords_matrix, fun = distHaversine)
  
  # remove diagonal and lower triangle (distance to self = 0)
  dist_values <- dist_matrix[upper.tri(dist_matrix)]
  
  # summary stats
  list(
    mean = mean(dist_values),
    median = median(dist_values),
    max = max(dist_values)
  )
}

# calculate for Copenhagen
cph_dispersion <- avg_pairwise_dist(cph_coords)

# calculate for Paris
paris_dispersion <- avg_pairwise_dist(paris_coords)

# combine results
dispersion_comparison <- data.frame(
  city = c("Copenhagen", "Paris"),
  mean_pairwise_distance_m = c(cph_dispersion$mean, paris_dispersion$mean),
  median_pairwise_distance_m = c(cph_dispersion$median, paris_dispersion$median),
  max_pairwise_distance_m = c(cph_dispersion$max, paris_dispersion$max)
)

print(dispersion_comparison)
```

The data shows that Paris landmarks are more widely dispersed than those in Copenhagen. Paris has a higher average pairwise distance between landmarks (≈ 2,229 m) compared to Copenhagen (≈ 1,682 m), and the median distance is also greater (2,113 m vs. 1,370 m). This suggests that Paris landmarks are spread over a broader area, potentially requiring more travel between points of interest. Interestingly, however, Copenhagen has a higher maximum pairwise distance (≈ 5,874 m vs. Paris’ 4,841 m), indicating that while Copenhagen’s landmarks are generally more clustered, there are a few that are significantly farther out.

# analyzing hotel to landmark pairs within 1,000 meters
```{r}
# set your distance threshold (e.g., 1000 meters)
max_distance <- 1000

# filter hotel-landmark pairs within max_distance
hotels_near_landmarks <- cph_hotel_landmark %>%
  filter(distance_m <= max_distance)

# count hotels per landmark
hotel_counts_per_landmark <- hotels_near_landmarks %>%
  group_by(cph_landmark_name = landmark) %>%
  summarise(hotel_count = n_distinct(hotel)) %>%
  arrange(desc(hotel_count))

# join counts back to landmark geometry
landmarks_with_counts <- cph_landmark_points %>%
  left_join(hotel_counts_per_landmark, by = c("cph_landmark_name" = "cph_landmark_name"))

# replace NA hotel counts with 0
landmarks_with_counts$hotel_count[is.na(landmarks_with_counts$hotel_count)] <- 0

# create color palette for dynamic red shading
pal <- colorNumeric(
  palette = c("#FFC8C8", "#FF8A8A", "#FF5252", "#CC1F1F", "#800000"),
  domain = landmarks_with_counts$hotel_count
)

# create leaflet map with shaded markers
leaflet(landmarks_with_counts) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = ~rescale(hotel_count, to = c(5, 20)),
    color = ~pal(hotel_count),
    fillColor = ~pal(hotel_count),
    fillOpacity = 0.8,
    stroke = FALSE,
    label = ~paste0(cph_landmark_name, ": ", hotel_count, " hotels")
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = landmarks_with_counts$hotel_count,
    title = "Hotels per Landmark",
    opacity = 1
  ) %>%
  addControl(
  html = "<div style='font-size:18px; font-weight:bold; padding:5px;'>Number of Hotels within 1,000 Meters of Landmark - Copenhagen</div>",
    position = "topright"
  )
```

```{r}

max_distance <- 1000

# filter hotel-landmark pairs within max_distance
hotels_near_landmarks <- paris_hotel_landmark %>%
  filter(distance_m <= max_distance)

# count hotels per landmark
hotel_counts_per_landmark <- hotels_near_landmarks %>%
  group_by(landmark) %>%
  summarise(hotel_count = n_distinct(hotel)) %>%
  arrange(desc(hotel_count))

# join counts back to landmark geometry (assuming landmark_points is an sf object with geometry and landmark_name)
landmarks_with_counts <- paris_landmark_points %>%
  left_join(hotel_counts_per_landmark, by = c("landmark_name" = "landmark"))

# replace NA hotel counts with 0
landmarks_with_counts$hotel_count[is.na(landmarks_with_counts$hotel_count)] <- 0

# create color palette for dynamic red shading (5 shades of red, darkest = most hotels)
pal <- colorNumeric(
  palette = c("#FFC8C8", "#FF8A8A", "#FF5252", "#CC1F1F", "#800000"),
  domain = landmarks_with_counts$hotel_count
)

# create leaflet map with shaded markers
leaflet(landmarks_with_counts) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = ~rescale(hotel_count, to = c(5, 20)),
    color = ~pal(hotel_count),
    fillColor = ~pal(hotel_count),
    fillOpacity = 0.8,
    stroke = FALSE,
    label = ~paste0(landmark_name, ": ", hotel_count, " hotels")
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = landmarks_with_counts$hotel_count,
    title = "Hotels per Landmark",
    opacity = 1
  ) %>%
  addControl(
  html = "<div style='font-size:18px; font-weight:bold; padding:5px;'>Number of Hotels within 1,000 Meters of Landmark - Paris</div>",
    position = "topright"
  )
```

# finding shortest route in copenhagen and paris to compare
```{r}
# find the shortest route in Copenhagen
shortest_cph <- cph_hotel_landmark %>%
  summarise(
    min_distance_m = min(distance_m, na.rm = TRUE),
    closest_hotel = hotel[which.min(distance_m)],
    closest_landmark = landmark[which.min(distance_m)]
  )

# find the shortest route in Paris
shortest_paris <- paris_hotel_landmark %>%
  summarise(
    min_distance_m = min(distance_m, na.rm = TRUE),
    closest_hotel = hotel[which.min(distance_m)],
    closest_landmark = landmark[which.min(distance_m)]
  )

shortest_cph
shortest_paris
```

# finding how many hotels are within 7km radius of all landmarks
```{r}
# define threshold
threshold <- 5000  # meters

# copenhagen
# Get number of unique landmarks
cph_n_landmarks <- cph_hotel_landmark %>% pull(landmark) %>% unique() %>% length()

# find hotels within 7 km of all landmarks
cph_hotels_within_all <- cph_hotel_landmark %>%
  filter(distance_m <= threshold) %>%
  group_by(hotel) %>%
  summarise(n_within_range = n_distinct(landmark)) %>%
  filter(n_within_range == cph_n_landmarks)

# view results
cph_hotels_within_all

# Paris
# get number of unique landmarks
prs_n_landmarks <- paris_hotel_landmark %>% pull(landmark) %>% unique() %>% length()

# find hotels within 7 km of all landmarks
prs_hotels_within_all <- paris_hotel_landmark %>%
  filter(distance_m <= threshold) %>%
  group_by(hotel) %>%
  summarise(n_within_range = n_distinct(landmark)) %>%
  filter(n_within_range == prs_n_landmarks)

# view results
prs_hotels_within_all

```

# hotels within 3 km of half the landmarks chosen at random
```{r}

set.seed(42)  # for reproducibility

# Copenhagen
# sample half the landmarks
c_landmarks_all <- unique(cph_hotel_landmark$landmark)
c_n_half <- floor(length(c_landmarks_all) / 2)
c_landmarks_sampled <- sample(c_landmarks_all, c_n_half)

# filter for only those sampled landmarks
cph_df_filtered <- cph_hotel_landmark %>%
  filter(landmark %in% c_landmarks_sampled)

# find hotels within 3 km of each sampled landmark
c_hotels_close <- cph_df_filtered %>%
  filter(distance_m <= 3000) %>%
  group_by(hotel) %>%
  summarise(n_within_range = n_distinct(landmark)) %>%
  filter(n_within_range >= c_n_half)

# view result
c_hotels_close

# Paris
# sample half the landmarks
p_landmarks_all <- unique(paris_hotel_landmark$landmark)
p_n_half <- floor(length(p_landmarks_all) / 2)
p_landmarks_sampled <- sample(p_landmarks_all, p_n_half)

# filter for only those sampled landmarks
paris_df_filtered <- paris_hotel_landmark %>%
  filter(landmark %in% p_landmarks_sampled)

# find hotels within 3 km of each sampled landmark
p_hotels_close <- paris_df_filtered %>%
  filter(distance_m <= 3000) %>%
  group_by(hotel) %>%
  summarise(n_within_range = n_distinct(landmark)) %>%
  filter(n_within_range >= p_n_half)

# view result
p_hotels_close

```